---
title: "Class_Assignment9.17"
author: "Adam Kaderbhai"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(modeldata)
```
# A Simple Bar chart showing the frequency of each specie included in the study
```{r}
#| message: false
#| warning: false
ggplot(scat, 
       aes(x = Species,
           fill = Species)) + 
  geom_bar(show.legend = FALSE) + 
  labs(title = "Frequency of Each Specie",
       x = "Species",
       y = "Frequency") +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal()
```
- `Bobcat` has the highest frequency, indicating it is the most frequently observed species in this dataset, with a frequency of approximately `60` observations.

- `Coyote` has the second-highest frequency, with around `28` observations.

- `Gray_fox` has the lowest frequency among the three species, with roughly `25` observations.
# A boxplot showing the Variation in the Taper Index `TI` of Each specie found in Different Locations
```{r}
#| message: false
#| warning: false
ggplot(scat, 
       aes(x = Location,
                 y = TI,
                 fill = Species)) +
  geom_boxplot(show.legend = FALSE) +
  facet_wrap(~Species) +
  scale_fill_brewer(palette = "Dark2") +
  theme_minimal()
```
- The median `TI` values are fairly similar across `Species` and `Location`, mostly between 1 and 2.

- `Bobcat` scats tend to have slightly higher `TI` at the `edge` compared to `middle` and `off_edge`.

- `Coyotes` show relatively consistent `TI` across locations, with a few outliers.

- `Gray_foxes` have more variability, especially at the `off_edge` location, and one extreme outlier with very high `TI`.
# A bar chart showing the relationship between Age and d15N Composition in the animal scat
```{r}
#| message: false
#| warning: false
ggplot(scat, 
       aes(x = Age,
           y = d15N,
           fill = Species)) + 
  geom_col(position = "dodge") + 
  labs(title = " ",
       x = "Age",
       y = "d15N") +
  scale_fill_brewer(palette = "Dark2") + theme_minimal()
```
`Coyotes` consistently have the highest `d15N` values at every age, often much higher than the other two species.

`Bobcats` tend to fall in the middle, with moderately high d15N values.

`Gray foxes` show the lowest d15N values overall, across all ages.