---
title: "ClassNotes9.19"
format: html
embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(modeldata)
```

## Arranging a data set by a variable

Let's arrange scat by `d15N`.

```{r}
scat %>%
  select(Species,
         Site,
         d15N) %>%
  arrange(desc(d15N)) %>%
  head() # show just the first few rows

```
We can arrange by multiple variables including categorical ones
```{r}
scat %>%
  select(Species,
         Site,
         d15N,
         Location) %>%
  arrange(Location,
          desc(d15N))

  
```
# Adding/modifying columns

Use the `mutate` command. Let's get the ratio of the diamter to `length` in the `scat` dataset.

```{r}
scat %>%
  select(Species,
         Diameter,
         Length) %>%
  mutate(Shape = Diameter / Length) %>%
  head()
```
We can also use this command to change an existing column 

Last Comment: `arrange` always gives back a set with the same dimensions as the original. `mutate` always has the same number of rows but possible more columns. `filter` has the same number of columns, and `select` has the same number of row.

