---
title: "Class notes template"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(readxl)
trout <- read_excel("trout.xlsx")
```
 
```{r}
glimpse(trout)
```

Do trout at the different locations have different hemoglobin levels?

$H_0$: the population means of `Hemo` are the same for all locations.

$H_a$: the population means of `Hemo` are not the same for all locations.

```{r}
trout <- trout %>% 
  mutate(Location = as.factor(Location))

ggplot(trout, aes(x = Location,
                  y = Hemo)) +
  geom_boxplot() +
  geom_jitter()
```

Let's run the test.

```{r}
model <- aov(Hemo ~ Location, 
             data = trout)
summary(model)
```

The tiny p-value indicates good evidence against the null. We can conclude that `Location` does give information about `Hemo`. Trout at different locations do seem to have different `Hemo` levels.

This does **not** say anything about which groups might be different. To do that, we need to do some pairwise comparisons. However, if we just run all of those directly, there is a high probability of getting at least one small p-value just by chance, even if there really isn't a difference.

## The Tukey post-hoc test

To correctly follow up a positive ANOVA, use a *Tukey Honest Standard Differences Test*, which does a pairwise comparison for each pair of groups, correcting for the multiple tests.

```{r}
TukeyHSD(model)
```

There is good evidence that 1 and 2 are different. The evidence that 1 and 3 are different is weaker, and the evidence that 1 and 4 are different is certainly insufficient.

## Example: penguins

$H_0$: bill length is the same for all three species.

$H_1$: bill length is not the same.

```{r}
ggplot(penguins, aes(x = bill_len,
                     y = species)) + 
  geom_boxplot()
```

```{r}
model <- aov(bill_len ~ species,
             data = penguins)
summary(model)
```

There is very strong evidence ($p < 4\times 10^{-16}$) that `bill_len` is not independent from `species`.
