---
title: "Class notes 9.8"
author: "Adam Kaderbhai"
date: "8 September 2025"
format: html
---

If you have `editor: visual` in your YAML, please delete it.

## Bar charts
Bar charts are used to visualize 1-2 categorical variables. We'll explore this using the `Carseats` data set from the `ISLR2` package.

```{r}
#| message: false
library(ISLR2)
library(tidyverse)
```

Use the data command to view all the datasets

```{r}
glimpse(Carseats)
```
Let's Consider `Shelvelov` and `US`

```{r}
ggplot(Carseats, aes(x = ShelveLoc,
                     fill = US )) + 
  geom_bar() + 
  scale_fill_brewer(palette = "Dark2")
```
This plot stresses the relative counts of the `ShelveLoc` variable. It also shows `US`, but relative proportions are not so clear. We can flip the variables to reverse this.

```{r}
ggplot(Carseats, aes(fill = ShelveLoc,
                     x = US )) + 
  geom_bar() + 
  scale_fill_brewer(palette = "Dark2")
```
Let's stress the relative proportions. Same data, same aesthetics.

```{r}
ggplot(Carseats, aes(fill = ShelveLoc,
                     x = US )) + 
  geom_bar(position = "fill") + 
  scale_fill_brewer(palette = "Dark2")
```

Check out the help file for more options. `?geom_bar`

```{r}
ggplot(Carseats, aes(fill = ShelveLoc,
                     x = US )) + 
  geom_bar(position = "dodge") + 
  scale_fill_brewer(palette = "Dark2")
```