---
title: "ClassNotes11.24"
author: Adam Kaderbhai
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
```

## Regression with a categorical explainer

```{r}
penguins_sm <- penguins %>% 
  filter(species != "Adelie")

ggplot(penguins_sm, aes(x = bill_dep,
                        y = bill_len,
                        color = species)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal()
```


Regression with a categorical explanatory variable 
- Well assume equal slopes for now
Roughly, we want one model for chinstrap and one model for Gentoo

Chinsteap -> y = mx+b
Gentoo -> y = mx + b

y = bill_len
x = bill_dep


More specifically y = {mx+b_chin
                       mx + b_chin + b_diff (for gentoo)}
                       
Let X_2 = O if chistrap and 1 if gentoo

y = b_



```{r}
model <- lm(bill_len ~ bill_dep + species,
            data = penguins_sm)
```
The form of this model is $bill_len = \beta_0 + \beta_1 * bill_dep * \beta_2 (isgentoo)$

```{r}
model
```

```{r}
library(broom)

penguins_aug <- augment(model)
glimpse()
```

