---
title: "Class notes 11.14"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(readxl)
parenting <- read_excel("parenting.xlsx")
```

## Linear regression

We're considering the relationship between 2 quantitative variables. For instance, is Danielle's grumpiness level related to the amount of sleep baby gets?

```{r}
glimpse(parenting)
```

Regression considers a possible *linear* relationship.

$H_0$: there is no linear relationship between the variables. The slop of the line of best fit is zero.

$H_a$: there is a linear relationship between the variable. The slope of the line of best fit is non-zero

A visualization is **essential** when doing a regerssion. You have to check that the relationship appears linear.

```{r}
ggplot(parenting, aes(x = baby.sleep,
                      y = dan.grump)) + 
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE)
```

We run the test with `lm`.

```{r}
model <- lm(dan.grump ~ baby.sleep, 
            data = parenting)
summary(model)
```

The model here is `dan.grump = 85.782 - 2.742 * baby.sleep`. The p-value of the slope is $8.45\times 10^{-10}$, indicating strong evidence against the null. We conclude that in the larger population (all days after the baby sleeps), Danielle's grumpiness level *is* impacted by the amount of sleep the baby got. 

