---
title: "HomeWork7"
author: "Adam Kaderbhai"
format: html
embed-resources: true
---

```{r}
#| message: false
library(tidyverse)
library(gt)
```


```{r}
#| message: false

attrition_small <- read_csv("attrition_small.csv")
```

# Problem 1

The company claims that 10% of employees report their work/life balance as “best,” 65% report it as “better,” 20% as “good,” and only 5% as “bad.” Is this claim reasonable in light of the sample data?

- $H_o$: The companies claim is true
- $H_a$: The companies claim is false
```{r}
Observed_Worklife_by_count <- attrition_small %>% 
  group_by(WorkLifeBalance) %>%
  summarize(
    count = n()
  ) 

gt(Observed_Worklife_by_count) %>%
  tab_header("Work/Life Status by count")


Observed <- c(27, 157, 49, 17)

Probs <- c(.10, .65, .20, .05)



chisq.test(Observed, p = Probs)
```

The companies claim would be reasonable if the sample data matched ours.


# Problem 2

The company claims that the education field of 55% of their employees is in the life science, 20% is in marketing, 20% is in technical, and 5% is other. Is this claim reasonable in light of the sample data?

- $H_o$: The companies claim is true
- $H_a$: The companies claim is false
```{r}
Education_field_by_count <- attrition_small %>% 
  group_by(EducationField) %>%
  summarize(
    count = n()
  ) 

gt(Education_field_by_count) %>% 
  tab_header("Education Field by Count")


observed <- c(150, 44, 22, 34)

probs <- c(.55, .20, .05, .20)

chisq.test(observed, p = probs)

```
The companies claim would be reasonable if the sample data matched ours.




