---
title: "HW 5.11"
author: "Adam Kaderbhai"
format: html
embed-resources: true
---

```{r}
#| message: false
library(tidyverse)
library(ISLR2)
library(gt)
```

# Problem 1
```{r}
mean_price <- mean(Carseats$Price)
mean_competitor_price <- mean(Carseats$CompPrice)
target_price <- 110

price_data <- data.frame(
  variable = c("mean_price", "mean_competitor_price", "target_price"),
  price = c(mean_price, mean_competitor_price, target_price)) 

gt(price_data)

ggplot(price_data, aes(x = variable,
                       y = price,
                       fill = variable)) +
  geom_col(show.legend = FALSE) +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Price Comparison", x = "Variable", y = "Value") +
  theme_minimal()
```
Hypothesis: The company's claim that the average price of their own seats is $110 is False.

Alternative: $H_a: \mu \ne 100$ 

Although the company's average price of carseats is very close from their claim of `$110`, the actual average price of their car seats is `$116` this is still lower than the average price of their competitors whose average price is `$125`.


# Problem 2

```{r}
Carseats$Price %>% 
  t.test(mu = 100)
```
The margin of error is 124.975 ± 1.261

# Problem 3
```{r}
penguins_table <- table(penguins$island, penguins$species)

addmargins(penguins_table)

```
# Problem 4

[My Github Repo](https://github.com/AdamKaderbhai/MATH250/blob/main/HomeWork6) ^[AdamKaderbhai Github Repo]


